---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# pruin

<!-- badges: start -->
<!-- badges: end -->

The goal of pruin is to compute ruin probability.

## Installation

You can install the development version of pruin from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("haydo1117/pruin")
```

R version 4.1+ is required.


## Example

### Computation of finite ruin probability for Poisson-Exponential process

Run `?ruin_prob_exp` for details. `n` is the number of terms used in the Gaver-Stehfest algorithm. 
Large `n` results in divergence due to numerical precision issue. Small `n` does not converge. 

```{r example}
library(pruin)

c <- 1
lambda <- 1
beta <- 1.2

ruin_prob_exp(u=1,t=1,c,lambda,beta)

## once `n` is chosen, switch off `try_n`
ruin_prob_exp(u=1,t=1,c,lambda,beta,n=7,try_n=FALSE)

```

### Computation of Inverse Laplace Transform in general

Run `?fn_gs` for details. 

```{r}
library(pruin)

## laplace transform of constant function k
Fconst <- function(x,k)k/x

fn_gs(Fconst,2,n=15,k=2,plot=FALSE)
```



