---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# pruin

<!-- badges: start -->
<!-- badges: end -->

The goal of pruin is to compute finite-time ruin probability via bivariate Laguerre series, see 
"Finite-time ruin probabilities using bivariate Laguerre series". Currently support 
(1) combination of exponentials, 
(2) exponential, 
(3) generalised inverse Gaussian, 
(4) Normal truncated above zero, and 
(5) Weibull.
For exponential distribution, Inverse Laplace Transform using Gaver-Stehfest algorithm is also included. 
As a by-product,
a function to perform Inverse Laplace Transform is also included.

## Installation

You can install the development version of pruin from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("haydo1117/pruin")
```

R version 4.1+ is required.


## Example

```{r}
library(pruin)
library(polynom)
```

Process parameters:
```{r}
c <- 1.1
lambda <- 1
```


This section displays the use of the package for different claim distributions. 
It aims to recover some of the values in tables in the article 
"Finite-time ruin probabilities using bivariate Laguerre series".
Run `?FUNCTION_NAME` for details of specific function.

Since `uscale_search` takes time, it is mostly commented out for this document. 
The specific chosen values of `u_scale` were determined using `uscale_search`.
Uncomment those lines for full results.

### Exponential distribution

Rate $\beta=1$ is used.

#### Inverse Laplace
Run `?ruin_prob_exp` for details. `n` is the number of terms used in the Gaver-Stehfest algorithm. 
Large `n` results in divergence due to numerical precision issue. Small `n` does not converge. 

```{r out.width="50%"}
ruin_prob_exp_gs(u=10,t=2,c,lambda,beta=1,plot=FALSE) # `plot=TRUE` for a visualisation

# once `n` is chosen, switch off `try_n`
ruin_prob_exp_gs(u=10,t=2,c,lambda,beta=1,n=7,try_n=FALSE)
```

#### Bivariate Laguerre

```{r}
exp_search <- uscale_search(u=10,t=2,c=c,lambda=lambda,family=exponential()(beta=1)) # search for u_scale
head(exp_search)

ruin_prob_ls(u=10,t=2,c,lambda,family=exponential()(beta=1),check=TRUE,u_scale = 0.6) # once u_scale is chosen
```

### Combination of exponentials

```{r}
ce1 <- comb_exponential()(w=c(2,-1),beta=c(1.5,3))
# uscale_search(u=10,t=2,c=1.1,lambda=1,family=ce1)
ruin_prob_ls(u=10,t=2,c=1.1,lambda=1,family=ce1,check=TRUE,u_scale = 1.0715193)

ce2 <- comb_exponential()(w=c(1/3,2/3),beta=c(0.5,2))
# uscale_search(u=10,t=2,c=1.1,lambda=1,family=ce2)
ruin_prob_ls(u=10,t=2,c=1.1,lambda=1,family=ce2,check=TRUE,u_scale = 0.6250552)
```

### Weibull

```{r}
w1 <- weibull()(alpha=2,beta=1/gamma(1+1/2))
# uscale_search(u=8,t=8,c=1.1,lambda=1,family=w1)
ruin_prob_ls(u=8,t=8,c=1.1,lambda=1,family=w1,check=TRUE,u_scale = 1.6982437)

w2 <- weibull()(alpha=3,beta=1/gamma(1+1/3))
# uscale_search(u=4,t=4,c=1.1,lambda=1,family=w2)
ruin_prob_ls(u=4,t=4,c=1.1,lambda=1,family=w2,check=TRUE,u_scale = 3.0199517)
```

### Generalised inverse Gaussian

```{r}
g1 <- gig()(a=5.03251,b=0.01987,alpha=2.5)
# uscale_search(u=4,t=4,c=1.1,lambda=1,family=g1)
ruin_prob_ls(u=4,t=4,c=1.1,lambda=1,family=g1,check=TRUE,u_scale = 2.6302680)

g2 <- gig()(a=0.32497,b=0.61543,alpha=-0.75)
gig_search <- uscale_search(u=4,t=4,c=1.1,lambda=1,family=g2) 
head(gig_search) # unstable ("err" too large)
```

### Normal truncated above zero

```{r}
tn1 <- truncated_normal()(mu=0,sigma=sqrt(pi/2))
# uscale_search(u=8,t=6,c=1.1,lambda=1,family=tn1)
ruin_prob_ls(u=8,t=6,c=1.1,lambda=1,family=tn1,check=TRUE,u_scale = 1.1481536)

tn_par <- 1/(1+dnorm(1)/(1-pnorm(-1)))
tn2 <- truncated_normal()(mu=tn_par,sigma=tn_par)
# uscale_search(u=6,t=4,c=1.1,lambda=1,family=tn2)
ruin_prob_ls(u=6,t=4,c=1.1,lambda=1,family=tn2,check=TRUE,u_scale = 1.4791084)
```


### Computation of Inverse Laplace Transform in general

```{r}
# laplace transform of constant function k
Fconst <- function(x,k)k/x
fn_gs(Fconst,2,n=15,k=2,plot=FALSE)
```



